---
title: "Analysis"
author: "Dionna Attinson"
date: "2/15/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(epitools)
library(viridis)
library(patchwork)
library(modelr)
library(MVN)
library(mgcv)
library(magrittr)
library(lavaan)
library(corrplot)
library(gmodels)
library(patchwork)
library(gridExtra)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
FinalBetancesData %>% 
linearregression1 = lm(PCS12 ~ HousingSatisfactionScore + total_PSS_score + Age + Income_Group + Race + Gender_Coded + Education_Group + Chronic_Disease, data = .,)
summary(linearregression1)

linearregression1 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)

```

##### Housing satisfaction was not signficantly associated with self-reported physical health score. For every year increase in age, physical health score decreased by 0.307 at the 5% level of signficance. Those who were in income group 2, had 4.621 higher physical health score, compared to those in income group 1, at a 5% level of signficance. 

```{r}
linearregression2 = lm(MCS12 ~ HousingSatisfactionScore + Age + Income_Group + Race + Gender_Coded + Education_Group + Chronic_Disease, data = FinalBetancesData)
summary(linearregression2)

linearregression2 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Housing satisfaction was signficantly associated with self-reported mental health score at the 5% level of significance. For every unit increase in Housing Satisfaction, mental health score increased by .99  at the 5% level of significance. In addition, those with a chronic disease's mental health score decreased by 8.91 in comparison to those without a chronic disease at a 5% level of significance. 