---
title: "Data Cleaning, Preliminary Analysis, and Final Data Analysis"
author: "Dionna Attinson"
date: "2/7/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(epitools)
library(imputeTS)
library(viridis)
library(patchwork)
library(modelr)
library(MVN)
library(mgcv)
library(magrittr)
library(lavaan)
library(corrplot)
library(gmodels)
library(patchwork)
library(gridExtra)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
SurveyData = 
  readxl::read_xls('./SFH_Survey_CODED_CLEANED.xls') %>% 
  janitor::clean_names()
```

### Data Cleaning and Recoding (Betances Data)
```{r}
SurveyData =
 SurveyData %>% 
 rename(Gender = x1_to_which_gender_identity_do_you_most_identify_selected_choice,
        Education = x3_what_is_the_highest_level_of_education_you_have_completed_selected_choice,
        Race = x4_which_best_describes_your_race_ethnicity_check_all_that_apply_selected_choice,
        Income = x7_hand_participant_income_card_which_of_these_groups_on_this_card_is_closest_to_your_total_personal_income_before_taxes_during_the_last_year_please_include_income_from_all_sources_your_salaries_wages_and_any_benefits_including_social_security_welfare_gifts_or_any_other_income_please_do_not_include_food_stamps_or_rental_subsidies_tell_me_the_code_for_the_amount_you_got_last_year_or_the_code_for_the_amount_you_usually_get_per_month_if_necessary_work_with_resident_to_construct_a_monthly_personal_and_household_income)
```

##### In the code chunk below, a number of variables were recoded as binary. 
* `Smoke` was the variable that identified the smoking status of an individual. Those who responded that they smoke ANY products including marijuana, THC, hash, hookah, E-cigarettes, Cigarettes or Cigarillos were coded as `1` = Smoker. Those who responded that they did not smoke any products were coded as `0`, Non-smoker.
* Evidence of mice/rats, cockroaches, and having a mildew odor present in the home were all coded as `1` = Yes, `0` = No.
```{r}
VariablesofInterest = 
  SurveyData %>% 
  select(age, Gender, Race, pss_5, pss_6, pss_8_108, pss_8_109, pss_10, sf36_1, sf36_4, 
        sf36_6, sf36_14, sf36_15, sf36_18, sf36_19, sf36_22, sf36_26, sf36_27, sf36_28,
        sf36_32, Education, Income, repairs_needed, asthma, anxiety, lungd, depression,
        heartdisease, cancer, diabetes,hypertension, products_used, neg_health_building, 
        micerats, cockroaches, mildew, homesat,buildsat, neighborhoodsat, propmansat, sumtempsat, wintempsat) %>% 
  
  rename(Smoke = products_used) %>% 
  rename(Age = age) %>% 
  mutate(Smoke = recode(Smoke, 
                    "Dont smoke" = "0",
                    "Cigarettes and/or Cigarillos (common brands are Marlboro, Black & Mild)" = "1",
                    "Hookah" = "1",
                    "Cigarettes and/or Cigarillos (common brands are Marlboro, Black & Mild) ,Marijuana, hash, THC, grass, pot, or weed" = "1",
                    "Cigarettes and/or Cigarillos (common brands are Marlboro, Black & Mild) ,E-cigarettes (look like regular cigarettes, but are battery-powered),Marijuana, hash, THC, grass, pot, or weed,Hookah" = "1",
                    "E-cigarettes (look like regular cigarettes, but are battery-powered)" = "1",
                   "Marijuana, hash, THC, grass, pot, or weed" = "1"),
        micerats = recode(micerats, 
                          "Yes" = "1", 
                          "No" = "0"),
        cockroaches = recode(cockroaches, 
                             "Yes" = "1", 
                             "No" = "0"),
        mildew = recode(mildew, 
                            "Yes" = "1",
                            "No" = "0"),
        neg_health_building = recode(neg_health_building, 
                                      "Yes" = "1",
                                      "No" = "0"),
        Race = recode(Race,
                      "Hispanic or Latino" = "Hispanic or Latinx",
                      "Other, specify:" = "Other",
                      "Hispanic or Latino,Native American/ Alaskan Native" = "Other",
                      "Non- Hispanic Black or African American,Asian or Pacific Islander" = "Other",
                      "Non- Hispanic Black or African American,Hispanic or Latino" = "Other",
                      "Non-Hispanic White,Non- Hispanic Black or African American,Hispanic or Latino,Native American/ Alaskan Native,Asian or Pacific Islander" = "Other"))
```

##### Below, I scored the Perceived Stress Scale (PSS-4) and the SF-12. The PSS coding instructions were as follows: code `0` - never, `1` - almost never, `2` - sometimes, `3` - fairly often, `4` - very often & reverse your scores for questions 4, 5, 7, and 8. On these 4 questions, change the scores to: 0 = 4, 1 = 3, 2 = 2, 3 = 1, 4 = 0. 

```{r}
##PSS Scale
CodedData_VariablesofInterest =
  VariablesofInterest %>% 
  mutate(
    pss_5_coded = pss_5,
    pss_6_coded = pss_6,
    pss_8_coded = pss_8_108,
    pss_9_coded = pss_8_109,
    pss_10_coded = pss_10, 
    gh1 = sf36_1,
    pf02 = sf36_4,  
    pf04 = sf36_6,
    rp2 = sf36_14,  
    rp3 = sf36_15,  
    re2 = sf36_18,
    re3 = sf36_19,
    bp2 = sf36_22,
    mh3 = sf36_26,
    vt2 = sf36_27,
    mh4 = sf36_28,
    sf2 = sf36_32) %>% 
    
  mutate(pss_5_coded = recode(pss_5_coded, 
                        "Never" = "4",
                        "Almost never" = "3",
                        "Sometimes" = "2", 
                        "Fairly often" = "1",
                        "Very often" = "0"), 
         pss_6_coded = recode(pss_6_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         pss_8_coded = recode(pss_8_coded, 
                        "Never" = "4",
                        "Almost never" = "3",
                        "Sometimes" = "2", 
                        "Fairly often" = "1",
                        "Very often" = "0"),
         pss_9_coded = recode(pss_9_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         pss_10_coded = recode(pss_10_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         gh1 = recode(gh1,
                         "Excellent" = "1",
                         "Very Good" = "2",
                         "Good" = "3",
                         "Fair" = "4",
                         "Poor" = "5"),
         pf02 = recode(pf02, 
                         "Yes, limited a lot" = "1",
                         "Yes, limited a little" = "2",
                         "No, not limited at all" = "3"),
         pf04 = recode(pf04, 
                         "Yes, limited a lot" = "1",
                         "Yes, limited a little" = "2",
                         "No, not limited at all" = "3"),
         rp2 = recode(rp2, 
                         "Yes" = "1",
                         "No" = "2"),
         rp3 = recode(rp3, 
                         "Yes" = "1",
                         "No" = "2"), 
         re2 = recode(re2, 
                         "Yes" = "1",
                         "No" = "2"), 
         re3 = recode(re3, 
                         "Yes" = "1",
                         "No" = "2"),
         bp2 = recode(bp2,
                         "Not at all" = "1",
                         "A little bit" = "2",
                         "Moderately" = "3",
                         "Quite a bit" = "4",
                         "Extremely" = "5"),
         mh3 =  recode(mh3,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         vt2 = recode(vt2,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         mh4 = recode(mh4,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         sf2 = recode(sf2,
                         "All of the time" = "1",
                         "Most of the time" = "2",
                         "Some of the time" = "3",
                         "A little of the time" = "4",
                         "None of the time" = "5")) 
```

##### The code chunk below now represents the second step of scoring the SF-12. This requires reversing the scores of certain questions so that a higher SF-12 score means better health.
```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
     gh1 = recode(gh1,
                         "1" = "5",
                         "2" = "4",
                         "3" = "3",
                         "4" = "2",
                         "5" = "1"),
         bp2 = recode(bp2,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"),
         mh3 = recode(mh3,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"),
         vt2 = recode(vt2,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"))
```

##### Next, I created dummy variables for each of the SF-12 variables.
```{r}
  CodedData_VariablesofInterest$pf02_1 <- as.numeric(CodedData_VariablesofInterest$pf02 == 1L) 
  CodedData_VariablesofInterest$pf02_2 <- as.numeric(CodedData_VariablesofInterest$pf02 == 2L) 
  CodedData_VariablesofInterest$pf04_1 <- as.numeric(CodedData_VariablesofInterest$pf04 == 1L) 
  CodedData_VariablesofInterest$pf04_2 <- as.numeric(CodedData_VariablesofInterest$pf04 == 2L) 
  CodedData_VariablesofInterest$rp2_1 <- as.numeric(CodedData_VariablesofInterest$rp2 == 1L) 
  CodedData_VariablesofInterest$rp3_1 <- as.numeric(CodedData_VariablesofInterest$rp3 == 1L) 
  CodedData_VariablesofInterest$bp2_1 <- as.numeric(CodedData_VariablesofInterest$bp2 == 1L) 
  CodedData_VariablesofInterest$bp2_2 <- as.numeric(CodedData_VariablesofInterest$bp2 == 2L) 
  CodedData_VariablesofInterest$bp2_3 <- as.numeric(CodedData_VariablesofInterest$bp2 == 3L) 
  CodedData_VariablesofInterest$bp2_4 <- as.numeric(CodedData_VariablesofInterest$bp2 == 4L) 
  CodedData_VariablesofInterest$gh1_1 <- as.numeric(CodedData_VariablesofInterest$gh1 == 1L) 
  CodedData_VariablesofInterest$gh1_2 <- as.numeric(CodedData_VariablesofInterest$gh1 == 2L) 
  CodedData_VariablesofInterest$gh1_3 <- as.numeric(CodedData_VariablesofInterest$gh1 == 3L) 
  CodedData_VariablesofInterest$gh1_4 <- as.numeric(CodedData_VariablesofInterest$gh1 == 4L) 
  CodedData_VariablesofInterest$vt2_1 <- as.numeric(CodedData_VariablesofInterest$vt2 == 1L) 
  CodedData_VariablesofInterest$vt2_2 <- as.numeric(CodedData_VariablesofInterest$vt2 == 2L) 
  CodedData_VariablesofInterest$vt2_3 <- as.numeric(CodedData_VariablesofInterest$vt2 == 3L) 
  CodedData_VariablesofInterest$vt2_4 <- as.numeric(CodedData_VariablesofInterest$vt2 == 4L) 
  CodedData_VariablesofInterest$vt2_5 <- as.numeric(CodedData_VariablesofInterest$vt2 == 5L) 
  CodedData_VariablesofInterest$sf2_1 <- as.numeric(CodedData_VariablesofInterest$sf2 == 1L) 
  CodedData_VariablesofInterest$sf2_2 <- as.numeric(CodedData_VariablesofInterest$sf2 == 2L) 
  CodedData_VariablesofInterest$sf2_3 <- as.numeric(CodedData_VariablesofInterest$sf2 == 3L) 
  CodedData_VariablesofInterest$sf2_4 <- as.numeric(CodedData_VariablesofInterest$sf2 == 4L) 
  CodedData_VariablesofInterest$re2_1 <- as.numeric(CodedData_VariablesofInterest$re2 == 1L) 
  CodedData_VariablesofInterest$re3_1 <- as.numeric(CodedData_VariablesofInterest$re3 == 1L) 
  CodedData_VariablesofInterest$mh3_1 <- as.numeric(CodedData_VariablesofInterest$mh3 == 1L) 
  CodedData_VariablesofInterest$mh3_2 <- as.numeric(CodedData_VariablesofInterest$mh3 == 2L) 
  CodedData_VariablesofInterest$mh3_3 <- as.numeric(CodedData_VariablesofInterest$mh3 == 3L) 
  CodedData_VariablesofInterest$mh3_4 <- as.numeric(CodedData_VariablesofInterest$mh3 == 4L) 
  CodedData_VariablesofInterest$mh3_5 <- as.numeric(CodedData_VariablesofInterest$mh3 == 5L) 
  CodedData_VariablesofInterest$mh4_1 <- as.numeric(CodedData_VariablesofInterest$mh4 == 1L) 
  CodedData_VariablesofInterest$mh4_2 <- as.numeric(CodedData_VariablesofInterest$mh4 == 2L) 
  CodedData_VariablesofInterest$mh4_3 <- as.numeric(CodedData_VariablesofInterest$mh4 == 3L) 
  CodedData_VariablesofInterest$mh4_4 <- as.numeric(CodedData_VariablesofInterest$mh4 == 4L) 
  CodedData_VariablesofInterest$mh4_5 <- as.numeric(CodedData_VariablesofInterest$mh4 == 5L) 
```  

##### I then created SF-12 weights and aggregates of indicator variables using physical and mental regression weights. 
```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(RAWPCS12 =
                   (-7.23216*pf02_1) + (-3.45555*pf02_2) +
                   (-6.24397*pf04_1) + (-2.73557*pf04_2) +
                   (-4.61617*rp2_1) + 
                   (-5.51747*rp3_1) +
                   (-11.25544*bp2_1) + (-8.38063*bp2_2) +
                   (-6.50522*bp2_3) + (-3.80130*bp2_4) + (-8.37399*gh1_1) +
                   (-5.56461*gh1_2) + (-3.02396*gh1_3) + (-1.31872*gh1_4) +
                   (-2.44706*vt2_1) + (-2.02168*vt2_2) + (-1.6185*vt2_3) +
                   (-1.14387*vt2_4) + (-0.42251*vt2_5) + (-0.33682*sf2_1) +
                   (-0.94342*sf2_2) + (-0.18043*sf2_3) + (0.11038*sf2_4) +
                   (3.04365*re2_1) + (2.32091*re3_1) + (3.46638*mh3_1) +
                   (2.90426*mh3_2) + (2.37241*mh3_3) + (1.36689*mh3_4) +
                   (0.66514*mh3_5) + (4.61446*mh4_1) + (3.41593*mh4_2) +
                   (2.34247*mh4_3) + (1.28044*mh4_4) + (0.41188*mh4_5),

           RAWMCS12 =
                   (3.93115*pf02_1) + (1.8684*pf02_2) +
                   (2.68282*pf04_1) + (1.43103*pf04_2) + (1.4406*rp2_1) +
                   (1.66968*rp3_1) + (1.48619*bp2_1) + (1.76691*bp2_2) +
                   (1.49384*bp2_3) + (0.90384*bp2_4) + (-1.71175*gh1_1) +
                   (-0.16891*gh1_2) + (0.03482*gh1_3) + (-0.06064*gh1_4) +
                   (-6.02409*vt2_1) + (-4.88962*vt2_2) + (-3.29805*vt2_3) +
                   (-1.65178*vt2_4) + (-0.92057*vt2_5) + (-6.29724*sf2_1) +
                   (-8.26066*sf2_2) + (-5.63286*sf2_3) + (-3.13896*sf2_4) +
                   (-6.82672*re2_1) + (-5.69921*re3_1) + (-10.19085*mh3_1) +
                   (-7.92717*mh3_2) + (-6.31121*mh3_3) + (-4.09842*mh3_4) +
                   (-1.94949*mh3_5) + (-16.15395*mh4_1) + (-10.77911*mh4_2) +
                   (-8.09914*mh4_3) + (-4.59055*mh4_4) + (-1.95934*mh4_5))
```

##### We then computed a norm-based standardization to PCS12 and MCS12.
```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
  PCS12 = (RAWPCS12 + 56.57706),
  MCS12 = (RAWMCS12 + 60.75781))
```

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
pss_5_coded =  as.numeric(pss_5_coded),
pss_6_coded =  as.numeric(pss_6_coded),
pss_8_coded = as.numeric(pss_8_coded),
pss_9_coded = as.numeric(pss_9_coded),
pss_10_coded = as.numeric(pss_10_coded))
```

##### For the SF-12 scale, a standardized SF-12 score was created for both physical and mental health for each participant. This resulted in the variables `PCS12` and `MCS12`. The higher the SF-12 scores, the better their overall health-related quality of life.

##### For the PSS scale, a total score was created from the questions which became the variable `total_PSS_score`. I then created stress categories for future categorical analysis: 
* Scores ranging from 0-6 would be considered low stress.
* Scores ranging from 7-13 ould be considered moderate stress.
* Scores ranging from 14-20 would be considered high perceived stress.

##### Further, several other variables were transformed. Age was transformed to a categorical variable, creating the age groups of 18-24, 25-44, 45-64, and 65+. Gender was transformed into a binary variable (0 = male, 1 = female), and a Housing Satisfaction Score was created. A higher score indicates higher overall housing satisfaction. This score was created based off of the totals of four questions: 
* The participant's satisfaction with their apartment
* The participant's satisfcation with their building 
* The participant's satisfcation with their neighbhorhood
* The participant's satisfaction with their property management 
* The participant's satisfcation with their apartment's temperature in the summer
* The participant's satisfaction with their apartment's temperature in the winter

##### Lastly, a binary chronic disease outcome was created. If a participant responded that they or someone in their home have the following conditions, they were coded as `1`: 
* Cancer
* Heart Disease
* Diabetes
* Hypertension
* Asthma
* Lung Disease 
* Anxiety
* Depression 

```{r}
CodedData_VariablesofInterest = 
CodedData_VariablesofInterest %>% 
  mutate(
    total_PSS_score = pss_5_coded + pss_6_coded + pss_8_coded + pss_9_coded + pss_10_coded,
    PSS_Category = total_PSS_score, 
    PSS_Category = 
          case_when(PSS_Category >= 14 ~ 'High Perceived Stress',
                    PSS_Category >= 7 ~ 'Moderate Stress',
                    PSS_Category >= 0  ~ 'Low Stress'),
    Age_Category = Age,
    Age_Category = 
          case_when(Age_Category >= 65  ~ '65+',
                    Age_Category >= 45  ~ '45-64',
                    Age_Category >= 25 ~ '25-44',
                    Age_Category >= 18 ~ '18-24'),
    Gender_Coded = Gender, 
    Gender_Coded = recode(Gender_Coded, 
                          "Male" = "0",
                          "Female" = "1",
                          `Other, please specify:` = "Null"),
    AptSat_Coded = homesat, 
    AptSat_Coded = recode(AptSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"), 
    BuildingSat_Coded = buildsat, 
    BuildingSat_Coded = recode(BuildingSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
    NeighborhoodSat_Coded = neighborhoodsat, 
    NeighborhoodSat_Coded = recode(NeighborhoodSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   PropertyManSat_Coded = propmansat,
   PropertyManSat_Coded = recode(PropertyManSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   WinTempSat_Coded = wintempsat, 
   WinTempSat_Coded = recode(WinTempSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"), 
   SumTempSat_Coded = sumtempsat, 
   SumTempSat_Coded = recode(SumTempSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   WinTempSat_Coded = as.numeric(WinTempSat_Coded),
   SumTempSat_Coded = as.numeric(SumTempSat_Coded),
   AptSat_Coded = as.numeric(AptSat_Coded), 
   BuildingSat_Coded = as.numeric(BuildingSat_Coded),
   NeighborhoodSat_Coded = as.numeric(NeighborhoodSat_Coded),
   PropertyManSat_Coded = as.numeric(PropertyManSat_Coded),
   asthma = as.numeric(asthma),
   heartdisease = as.numeric(heartdisease),
   lungd = as.numeric(lungd),
   anxiety = as.numeric(anxiety),
   depression = as.numeric(depression),
   cancer = as.numeric(cancer),
   hypertension = as.numeric(hypertension),
   diabetes = as.numeric(diabetes),
   Chronic_Disease = asthma + heartdisease + lungd + anxiety + depression + cancer + hypertension + diabetes,
   Chronic_Disease = 
          case_when(Chronic_Disease >= 1 ~ '1',
                    Chronic_Disease == 0 ~ '0'))
```

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
  PropertyManSat_Coded = na_mean(PropertyManSat_Coded),
  WinTempSat_Coded = na_mean(WinTempSat_Coded),
  HousingSatisfactionScore = AptSat_Coded + BuildingSat_Coded + NeighborhoodSat_Coded + PropertyManSat_Coded + WinTempSat_Coded + SumTempSat_Coded,
  Satisfaction_Category = HousingSatisfactionScore,
  Satisfaction_Category = 
          case_when(Satisfaction_Category >= 18 ~ 'High Satisfaction',
                    Satisfaction_Category >= 12 ~ 'Moderate Satisfaction',
                    Satisfaction_Category >= 6  ~ 'Low Satisfaction'))
```

##### Below, I changed variables into factors and numeric values.

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
    Education = as.factor(Education),
    Gender_Coded = factor(Gender_Coded, levels = c("0","1")),
    Race = as.factor(Race),
    Age = as.numeric(Age),
    Income = factor(Income, levels = c("1", "2", "3", "4", "5", "6","7","8","9","10")),
    Smoke = factor(Smoke, levels = c("0", "1")),
    asthma = factor(asthma, levels = c("0","1")),
    lungd = factor(lungd, levels = c("0","1")),
    heartdisease = factor(heartdisease, levels = c("0","1")),
    anxiety = factor(anxiety, levels = c("0","1")),
    depression = factor(depression, levels = c("0","1")),
    cancer = factor(cancer,levels = c("0","1")),
    diabetes = factor(diabetes,levels = c("0","1")),
    micerats = factor(micerats,levels = c("0","1")),
    mildew = factor(mildew,levels = c("0","1")),
    cockroaches = factor(cockroaches,levels = c("0","1")),
    Chronic_Disease = factor(Chronic_Disease,levels = c("0","1"))
  )
```

##### In order to collapse education categories into fewer groups, education levels were finalized into three categories: 
* Less than high school (included those who never attended school, or did not complete a high school degree/G.E.D)
* High school/G.E.D
* Post High School Education (included those who attended vocational school, some college, 2-year/4-year college degree, and post graduate education)
```{r}
CodedData_VariablesofInterest = 
CodedData_VariablesofInterest %>% 
  mutate(Education = as.character(Education))
```

```{r}
CodedData_VariablesofInterest[99, "Education"] = "Never Attended School"
CodedData_VariablesofInterest[33, "Education"] = "Some College"
CodedData_VariablesofInterest[20, "Education"] = "Some College"
CodedData_VariablesofInterest[53, "Education"] = NA
```

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
  Education_Group = Education, 
  Education_Group = recode(Education_Group, 
                    "Never Attended School" = "Less than High School",
                    "Vocational School" = "Post High School Education",
                    "Some College" = "Post High School Education",
                    "2 Year Community College Degree" = "Post High School Education",
                    "4 Year College Degree" = "Post High School Education",
                    "Post-Graduate Degree" = "Post High School Education",
                    "High School/Secondary School Diploma" = 	"High School/G.E.D.",
                    "G.E.D." = "High School/G.E.D."))
```

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(Education_Group = as.factor(Education_Group))
```

##### In order to collapse income categories into fewer groups, education levels were finalized into three categories:
* `1` - Total personal income less than $10,000/year
* `2` - Total personal income between $10,000 - $24,000/year
* `3` - Total personal income $25,000k + /year

```{r}
CodedData_VariablesofInterest = 
  CodedData_VariablesofInterest %>% 
  mutate(
  Income_Group = Income, 
  Income_Group = recode(Income_Group, 
                    "2" = "1",
                    "3" = "1",
                    "4" = "2",
                    "5" = "2",
                    "6" = "3",
                    "7" = "3",
                    "8" = "3",
                    "9" = "3",
                    "10" = "3"),
  Income_Group = factor(Income_Group,levels= c("1","2","3")))
```

### Univariable/Descriptive Analyses
##### The tables below provide descriptive analyses of the independent and dependent variables. 

#### Outcome Variables 
##### Overall Physical Health Status (Outcome 1)
```{r}
CodedData_VariablesofInterest %>% 
  select(PCS12) %>% 
  summary() %>% 
  knitr::kable(col.names = c("Summary of Self-Reported Physical Health"))
```

```{r}
CodedData_VariablesofInterest %>% 
  drop_na() %>% 
  group_by(PCS12, Age_Category, Gender) %>% 
  summarize(n = n()) %>% 
    ggplot(aes(x= Age_Category, y = PCS12)) + geom_boxplot() +
  geom_jitter() +  facet_grid(. ~ Gender)
```

##### Overall Mentall Health Status (Outcome 2)
```{r}
CodedData_VariablesofInterest %>% 
  select(MCS12) %>% 
  summary() %>% 
  knitr::kable(col.names = c("Summary of Self-Reported Mental Health"))
```

```{r}
CodedData_VariablesofInterest %>% 
  drop_na() %>% 
  group_by(MCS12, Age_Category, Gender) %>% 
  summarize(n = n()) %>% 
    ggplot(aes(x= Age_Category, y = MCS12)) + geom_boxplot() +
  geom_jitter() +  facet_grid(. ~ Gender)
```

#### Independent Variables
```{r}
CodedData_VariablesofInterest %>% 
  group_by(Smoke) %>% 
  summarize(n=n()) %>% 
  mutate(
    percent = n/124*100) %>% 
  mutate(
    Smoke = recode(Smoke,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  knitr::kable(col.names=c("Smoking Status", "n", "Percent(%)"), digits = 2)
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Race) %>% 
  summarize(n=n()) %>%
   mutate(
  percent = n/124*100) %>% 
  mutate(
    Race = recode(Race,
                  "Other, specify:" = "Other")
  ) %>% 
  knitr::kable(col.names=c("Race/Ethnicity", "n", "Percent(%)"), digits = 2)
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Age_Category) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Age Group", "n", "Percent(%)"), digits = 2)
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Gender) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Gender", "n", "Percent(%)"), digits = 2)
```


```{r}
CodedData_VariablesofInterest %>% 
  drop_na() %>% 
  group_by(Age, Gender) %>% 
  ggplot(aes(x=Gender, y=Age)) + geom_violin() +
  labs(x = "Gender",
      y = "Age",
      title = "Figure 1. Distribution of Age by Gender")
```

```{r}
CodedData_VariablesofInterest %>% 
  drop_na() %>% 
  select(Gender, Age) %>% 
  group_by(Gender) %>% 
  summarize(Mean = mean(Age)) %>% 
  knitr::kable(col.names=c("Gender", "Mean Age"), digits = 2)
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(PSS_Category) %>% 
  summarize(n=n()) %>%
   mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Perceived Stress Score", "n", "Percent(%)"), digits = 2)

```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Income) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Income Category", "n", "Percent(%)"), digits = 2)
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Education) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Education Completed", "n", "Percent(%)"), digits = 2)
```

### 2x2 Tables and Visualizations
```{r}
CodedData_VariablesofInterest %>% 
  group_by(Age_Category, Gender) %>% 
  summarize(n=n()) %>%
  pivot_wider(
    names_from = Gender,
    values_from = n) %>% 
  knitr::kable(col.names = c("Age Group","Female (n)", "Male (n)", "Other (n)"))
```

```{r}
CodedData_VariablesofInterest %>%
  drop_na() %>% 
  group_by(Age_Category, Race) %>% 
  summarize(n=n()) %>%
  pivot_wider(
    names_from = Race,
    values_from = n
  ) %>% 
  knitr::kable(col.names = c("Age Group", "Hispanic or Latinx (n)", "Non-Hispanic Black or African American (n)", "Other (n)"))
```

```{r}
CodedData_VariablesofInterest %>%
  group_by(Race, Smoke) %>% 
  summarize(n=n()) %>%
  mutate(
    Smoke = recode(Smoke,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  pivot_wider(
    names_from = Race,
    values_from = n) %>% 
  knitr::kable(col.names = c("Smoking Status", "Hispanic or Latinx (n)", "Black or African American (n)", "Other (n)"))
```

```{r}
CodedData_VariablesofInterest %>%
  group_by(Age_Category, Smoke) %>% 
  drop_na() %>% 
  summarize(n=n()) %>%
  mutate(
    Smoke = recode(Smoke,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  pivot_wider(
    names_from = Age_Category,
    values_from = n) %>% 
  knitr::kable(col.names = c("Smoking Status", "18-24 (n)", "25-44 (n)", "45-64 (n)", "65+ (n)"))
```

```{r}
CodedData_VariablesofInterest %>% 
  mutate(
    Smoke = recode(Smoke,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")
  ) %>% 
ggplot(aes(x=Smoke, y=Age)) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0)) + 
  labs(x = "Smoking Status",
      y = "Age",
      title = "Figure 4. Distribution of Age by Smoking Status")
```

```{r}
CodedData_VariablesofInterest %>% 
  group_by(Age_Category, Education_Group) %>% 
  drop_na() %>% 
  ggplot(aes(x= Age_Category, fill = Education_Group)) + geom_bar() + coord_flip() +
  labs(x = "Age Group",
      y = "Count",
      color = "Education",
      title = "Figure 5. Education by Age Group") +
theme(legend.position = "bottom")
```

### Exploratory Analyses of Relationships between Variables
```{r}
CodedData_VariablesofInterest %>%  
  group_by(total_PSS_score, PCS12, Age_Category) %>%
  summarize(n=n()) %>% 
  ggplot(aes(x = total_PSS_score, y = PCS12,  color = Age_Category)) +
  geom_point(size = 1.5, alpha = 0.8) +
  geom_smooth(method =lm, color = "black", linetype = 1) +
  labs(
      x = "Perceived Stress Score",
      y = "Overall Health Score",
      color = "Age",
      title = "Figure 6. Relationship between Perceived Stress and Physical Health Status") +
      theme(legend.position = "right") 
```

```{r}
CodedData_VariablesofInterest %>%  
  group_by(HousingSatisfactionScore, PCS12, Age_Category) %>%
  summarize(n=n()) %>% 
  ggplot(aes(x = HousingSatisfactionScore, y = PCS12,  color = Age_Category)) +
  geom_point(size = 1, alpha = 0.5) +
  geom_smooth(method =lm, color = "black", linetype = 1) +
  labs(
      x = "Housing Satisfaction Score",
      y = "Overall Health Score",
      color = "Age",
      title = "Figure 7. Relationship between Housing Satisfaction and Physical Health Status") +
      theme(legend.position = "right")
```

```{r}
CodedData_VariablesofInterest %>%  
  group_by(total_PSS_score, MCS12, Age_Category) %>%
  summarize(n=n()) %>% 
  ggplot(aes(x = total_PSS_score, y = MCS12,  color = Age_Category)) +
  geom_point(size = 1.5, alpha = 0.8) +
  geom_smooth(method =lm, color = "black", linetype = 1) +
  labs(
      x = "Perceived Stress Score",
      y = "Overall Mental Health Score",
      color = "Age",
      title = "Figure 8. Relationship between Perceived Stress and Mental Health Status") +
      theme(legend.position = "right") 
```

```{r}
CodedData_VariablesofInterest %>%  
  group_by(HousingSatisfactionScore, MCS12, Age_Category) %>%
  summarize(n=n()) %>% 
  ggplot(aes(x = HousingSatisfactionScore, y = MCS12,  color = Age_Category)) +
  geom_point(size = 1, alpha = 0.5) +
  geom_smooth(method =lm, color = "black", linetype = 1) +
  labs(
      x = "Housing Satisfaction Score",
      y = "Overall Mental Health Score",
      color = "Age",
      title = "Figure 9. Relationship between Housing Satisfaction and Mental Health Status") +
      theme(legend.position = "right")
```

```{r}
CodedData_VariablesofInterest %>%  
  group_by(HousingSatisfactionScore, total_PSS_score, Age_Category) %>%
  summarize(n=n()) %>% 
  ggplot(aes(x = HousingSatisfactionScore, y = total_PSS_score,  color = Age_Category)) +
  geom_point(size = 1.5, alpha = 0.8) +
  geom_smooth(method =lm, color = "black", linetype = 1) +
  labs(
      x = "Housing Satisfaction Score",
      y = "Perceived Stress Score",
      color = "Age",
      title = "Figure 10. Relationship between Housing Satisfaction and Perceived Stress") +
      theme(legend.position = "right")
```

### Initial bivariate analyses 

##### Bivariate model 1 : PCS12 ~ b_0 + b_1 total_PSS_score_i 
```{r}
bivariate1 = lm(PCS12 ~ total_PSS_score, data = CodedData_VariablesofInterest)
summary(bivariate1) 

bivariate1 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 2 : PCS12 ~ b_0 + b_1 Age_i 
```{r}
bivariate2 = lm(PCS12 ~ Age, data = CodedData_VariablesofInterest)
summary(bivariate2)

bivariate2 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 3 : PCS12 ~ b_0 + b_1 Race_i 
```{r}
bivariate3 = lm(PCS12 ~ Race, data = CodedData_VariablesofInterest)
summary(bivariate3)

bivariate3 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 4 : PCS12 ~ b_0 + b_1 Income_i 
```{r}
bivariate4 = lm(PCS12 ~ Income_Group, data = CodedData_VariablesofInterest)
summary(bivariate4)

bivariate4 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 5 : PCS12 ~ b_0 + b_1 Smoke_i 
```{r}
bivariate5 = lm(PCS12 ~ Smoke, data = CodedData_VariablesofInterest)
summary(bivariate5)

bivariate5 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 6 : PCS12 ~ b_0 + b_1 ChronicDisease_i 
```{r}
bivariate6 = lm(PCS12 ~ Chronic_Disease, data = CodedData_VariablesofInterest)
summary(bivariate6)

bivariate6 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 7 : PCS12 ~ b_0 + b_1 TotalSatisfactionScore_i 
```{r}
bivariate7 = lm(PCS12 ~ HousingSatisfactionScore, data = CodedData_VariablesofInterest)
summary(bivariate7)

bivariate7 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 7a : PCS12 ~ b_0 + b_1 Education_i 

```{r}
bivariate7a = lm(PCS12 ~ Education_Group, data = CodedData_VariablesofInterest)
summary(bivariate7a)

bivariate7a %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 8 : MCS12 ~ b_0 + b_1 TotalSatisfactionScore_i 
```{r}
bivariate8 = lm(MCS12 ~ HousingSatisfactionScore, data = CodedData_VariablesofInterest)
summary(bivariate8)

bivariate8 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 9 : MCS12 ~ b_0 + b_1 Smoke_i 
```{r}
bivariate9 = lm(MCS12 ~ Smoke, data = CodedData_VariablesofInterest)
summary(bivariate9)

bivariate9 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 10 : MCS12 ~ b_0 + b_1 ChronicDisease_i 
```{r}
bivariate10 = lm(MCS12 ~ Chronic_Disease, data = CodedData_VariablesofInterest)
summary(bivariate10)

bivariate10 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 11 : MCS12 ~ b_0 + b_1 Income_i 
```{r}
bivariate11 = lm(MCS12 ~ Income_Group, data = CodedData_VariablesofInterest)
summary(bivariate11)

bivariate11 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 12 : MCS12 ~ b_0 + b_1 Race_i 
```{r}
bivariate12 = lm(MCS12 ~ Race, data = CodedData_VariablesofInterest)
summary(bivariate12) 

bivariate12 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 13 : MCS12 ~ b_0 + b_1 Age_i
```{r}
bivariate13 = lm(MCS12 ~ Age, data = CodedData_VariablesofInterest)
summary(bivariate13)

bivariate13 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 14 : MCS12 ~ b_0 + b_1 Gender_i 
```{r}
bivariate14 = lm(MCS12 ~ Gender_Coded, data = CodedData_VariablesofInterest)
summary(bivariate14)

bivariate14 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 15 : PCS12 ~ b_0 + b_1 Gender_i 
```{r}
bivariate15 = lm(PCS12 ~ Gender_Coded, data = CodedData_VariablesofInterest)
summary(bivariate15)

bivariate15 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 16 : MCS12 ~ b_0 + b_1 Education_i 
```{r}
bivariate16 = lm(MCS12 ~ Education_Group, data = CodedData_VariablesofInterest)
summary(bivariate16)

bivariate16 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Bivariate model 17 : MCS12 ~ b_0 + b_1 total_PSS_score_i 
```{r}
bivariate17 = lm(MCS12 ~ total_PSS_score, data = CodedData_VariablesofInterest)
summary(bivariate17)

bivariate17 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

### Checking for multicollinearity
```{r}
Correlation_Matrix = 
CodedData_VariablesofInterest %>% 
  select(HousingSatisfactionScore, total_PSS_score, Age, MCS12, PCS12) 
```

```{r}
CM1 = cor(Correlation_Matrix, use = "complete.obs")
round(CM1,2)
corrplot(CM1, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```

### Final dataset
```{r}
FinalBetancesData = 
  CodedData_VariablesofInterest %>% 
  select(Age, Age_Category, Gender, Gender_Coded, Education_Group, Income_Group, Race, Smoke, Chronic_Disease, HousingSatisfactionScore, MCS12, PCS12, total_PSS_score) 

FinalBetancesData %>% 
    rename(
    Smoking_Status = Smoke,
    Age_Group = Age_Category 
  )
```

### Final Analysis (Regression Models)

```{r}
linearregression1 = lm(PCS12 ~ HousingSatisfactionScore + total_PSS_score + Age + Income_Group + Race + Gender_Coded + Education_Group + Chronic_Disease, data = FinalBetancesData)
summary(linearregression1)

linearregression1 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)

```

##### Housing satisfaction was not signficantly associated with self-reported physical health score. For every year increase in age, physical health score decreased by 0.307 at the 5% level of signficance. Those who were in income group 2, had 4.621 higher physical health score, compared to those in income group 1, at a 5% level of signficance. 

```{r}
linearregression2 = lm(MCS12 ~ HousingSatisfactionScore + Age + Income_Group + Race + Gender_Coded + Education_Group + Chronic_Disease, data = FinalBetancesData)
summary(linearregression2)

linearregression2 %>% 
  broom::tidy() %>% 
  mutate(
         High_CI = estimate + 1.96*std.error,
         Low_CI = estimate - 1.96*std.error) %>% 
  select(term, estimate, p.value, Low_CI, High_CI) %>% 
  knitr::kable(digits = 3)
```

##### Housing satisfaction was signficantly associated with self-reported mental health score at the 5% level of significance. For every one unit increase in Housing Satisfaction, mental health score increased by .99  at the 5% level of significance. In addition, those with a chronic disease's mental health score was 8.91 lower compared to those without a chronic disease at a 5% level of significance. 

```{r}
estimates_full = tibble(
  outcome = c("Physical Health Score", "Mental Health Score"),
  Estimate = c(0.335, .990),
  pvalue = c(0.310, 0.002),
  Low_CI = c(-0.308, 0.376),
  High_CI = c(0.979, 1.605))
estimates_full %>% 
  ggplot(aes(x = Estimate, y = outcome, group = 1)) +
  geom_vline(aes(xintercept = 0), linetype = "dashed") +
  geom_errorbarh(aes(xmax = High_CI, xmin = Low_CI), size = 0.5, height = 0.2) +
  geom_point(size = 3.5, color = "orange") +
  ylab("") +
  xlab("Estimate") +
  ggtitle("Housing Satisfaction and estimates of select outcomes") +
  labs(
    caption = "Data from Smoke-free Housing Study Betances Site (2019)"
  )
```

### Data Cleaning and Recoding (SoBRO Data)
```{r}
SoBROData = 
  readxl::read_xls('./SOBRODATA.xls') %>% 
  janitor::clean_names()
```

##### In the code chunk below, a number of variables were recoded as binary. 
* `Smoke` was the variable that identified the smoking status of an individual. Those who responded that they smoke ANY products including marijuana, THC, hash, hookah, E-cigarettes, Cigarettes or Cigarillos were coded as `1` = Smoker. Those who responded that they did not smoke any products were coded as `0`, Non-smoker.
* Evidence of mice/rats, cockroaches, and having a mildew odor present in the home were all coded as `1` = Yes, `0` = No.
```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
        micerats = recode(micerats, 
                          "Yes" = "1", 
                          "No" = "0"),
        cockroach = recode(cockroach, 
                             "Yes" = "1", 
                             "No" = "0"),
        mildew = recode(mildew, 
                            "Yes" = "1",
                            "No" = "0"),
        buildingillhealth = recode(buildingillhealth, 
                                      "Yes" = "1",
                                      "No" = "0"),
        race = recode(race,
                      "Hispanic or Latino" = "Hispanic or Latinx",
                      "Other, specify:" = "Other"))
```

##### Below, I scored the Perceived Stress Scale (PSS-4) and the SF-12. The PSS coding instructions were as follows: code `0` - never, `1` - almost never, `2` - sometimes, `3` - fairly often, `4` - very often & reverse your scores for questions 4, 5, 7, and 8. On these 4 questions, change the scores to: 0 = 4, 1 = 3, 2 = 2, 3 = 1, 4 = 0. 

```{r}
##PSS Scale
SoBROData = 
  SoBROData %>% 
  mutate(
    pss_1_coded = pss1,
    pss_2_coded = pss2,
    pss_3_coded = pss3,
    pss_4_coded = pss4,
    pss_5_coded = pss5) %>% 
    
  mutate(pss_1_coded = recode(pss_1_coded, 
                        "Never" = "4",
                        "Almost never" = "3",
                        "Sometimes" = "2", 
                        "Fairly often" = "1",
                        "Very often" = "0"), 
         pss_2_coded = recode(pss_2_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         pss_3_coded = recode(pss_3_coded, 
                        "Never" = "4",
                        "Almost never" = "3",
                        "Sometimes" = "2", 
                        "Fairly often" = "1",
                        "Very often" = "0"),
         pss_4_coded = recode(pss_4_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         pss_5_coded = recode(pss_5_coded, 
                        "Never" = "0",
                        "Almost never" = "1",
                        "Sometimes" = "2", 
                        "Fairly often" = "3",
                        "Very often" = "4"),
         gh1 = recode(gh1,
                         "Excellent" = "1",
                         "Very Good" = "2",
                         "Good" = "3",
                         "Fair" = "4",
                         "Poor" = "5"),
         pf02 = recode(pf02, 
                         "Yes, limited a lot" = "1",
                         "Yes, limited a little" = "2",
                         "No, not limited at all" = "3"),
         pf04 = recode(pf04, 
                         "Yes, limited a lot" = "1",
                         "Yes, limited a little" = "2",
                         "No, not limited at all" = "3"),
         rp2 = recode(rp2, 
                         "Yes" = "1",
                         "No" = "2"),
         rp3 = recode(rp3, 
                         "Yes" = "1",
                         "No" = "2"), 
         re2 = recode(re2, 
                         "Yes" = "1",
                         "No" = "2"), 
         re3 = recode(re3, 
                         "Yes" = "1",
                         "No" = "2"),
         bp2 = recode(bp2,
                         "Not at all" = "1",
                         "A little bit" = "2",
                         "Moderately" = "3",
                         "Quite a bit" = "4",
                         "Extremely" = "5"),
         mh3 =  recode(mh3,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         vt2 = recode(vt2,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         mh4 = recode(mh4,
                          "All of the time" = "1",
                          "Most of the time" = "2",
                          "A good bit of the time" = "3",
                          "Some of the time" = "4",
                          "A little of the time" = "5",
                          "None of the time" = "6"),
         sf2 = recode(sf2,
                         "All of the time" = "1",
                         "Most of the time" = "2",
                         "Some of the time" = "3",
                         "A little of the time" = "4",
                         "None of the time" = "5")) 
```

##### The code chunk below now represents the second step of scoring the SF-12. This requires reversing the scores of certain questions so that a higher SF-12 score means better health.
```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
     gh1 = recode(gh1,
                         "1" = "5",
                         "2" = "4",
                         "3" = "3",
                         "4" = "2",
                         "5" = "1"),
         bp2 = recode(bp2,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"),
         mh3 = recode(mh3,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"),
         vt2 = recode(vt2,
                          "1" = "6",
                          "2" = "5",
                          "3" = "4",
                          "4" = "3",
                          "5" = "2",
                          "6" = "1"))
```

##### Next, I created dummy variables for each of the SF-12 variables.
```{r}
  SoBROData$pf02_1 <- as.numeric(SoBROData$pf02 == 1L) 
  SoBROData$pf02_2 <- as.numeric(SoBROData$pf02 == 2L) 
  SoBROData$pf04_1 <- as.numeric(SoBROData$pf04 == 1L) 
  SoBROData$pf04_2 <- as.numeric(SoBROData$pf04 == 2L) 
  SoBROData$rp2_1 <- as.numeric(SoBROData$rp2 == 1L) 
  SoBROData$rp3_1 <- as.numeric(SoBROData$rp3 == 1L) 
  SoBROData$bp2_1 <- as.numeric(SoBROData$bp2 == 1L) 
  SoBROData$bp2_2 <- as.numeric(SoBROData$bp2 == 2L) 
  SoBROData$bp2_3 <- as.numeric(SoBROData$bp2 == 3L) 
  SoBROData$bp2_4 <- as.numeric(SoBROData$bp2 == 4L) 
  SoBROData$gh1_1 <- as.numeric(SoBROData$gh1 == 1L) 
  SoBROData$gh1_2 <- as.numeric(SoBROData$gh1 == 2L) 
  SoBROData$gh1_3 <- as.numeric(SoBROData$gh1 == 3L) 
  SoBROData$gh1_4 <- as.numeric(SoBROData$gh1 == 4L) 
  SoBROData$vt2_1 <- as.numeric(SoBROData$vt2 == 1L) 
  SoBROData$vt2_2 <- as.numeric(SoBROData$vt2 == 2L) 
  SoBROData$vt2_3 <- as.numeric(SoBROData$vt2 == 3L) 
  SoBROData$vt2_4 <- as.numeric(SoBROData$vt2 == 4L) 
  SoBROData$vt2_5 <- as.numeric(SoBROData$vt2 == 5L) 
  SoBROData$sf2_1 <- as.numeric(SoBROData$sf2 == 1L) 
  SoBROData$sf2_2 <- as.numeric(SoBROData$sf2 == 2L) 
  SoBROData$sf2_3 <- as.numeric(SoBROData$sf2 == 3L) 
  SoBROData$sf2_4 <- as.numeric(SoBROData$sf2 == 4L) 
  SoBROData$re2_1 <- as.numeric(SoBROData$re2 == 1L) 
  SoBROData$re3_1 <- as.numeric(SoBROData$re3 == 1L) 
  SoBROData$mh3_1 <- as.numeric(SoBROData$mh3 == 1L) 
  SoBROData$mh3_2 <- as.numeric(SoBROData$mh3 == 2L) 
  SoBROData$mh3_3 <- as.numeric(SoBROData$mh3 == 3L) 
  SoBROData$mh3_4 <- as.numeric(SoBROData$mh3 == 4L) 
  SoBROData$mh3_5 <- as.numeric(SoBROData$mh3 == 5L) 
  SoBROData$mh4_1 <- as.numeric(SoBROData$mh4 == 1L) 
  SoBROData$mh4_2 <- as.numeric(SoBROData$mh4 == 2L) 
  SoBROData$mh4_3 <- as.numeric(SoBROData$mh4 == 3L) 
  SoBROData$mh4_4 <- as.numeric(SoBROData$mh4 == 4L) 
  SoBROData$mh4_5 <- as.numeric(SoBROData$mh4 == 5L) 
```  

##### I then created SF-12 weights and aggregates of indicator variables using physical and mental regression weights. 
```{r}
SoBROData = 
  SoBROData %>% 
  mutate(RAWPCS12 =
                   (-7.23216*pf02_1) + (-3.45555*pf02_2) +
                   (-6.24397*pf04_1) + (-2.73557*pf04_2) +
                   (-4.61617*rp2_1) + 
                   (-5.51747*rp3_1) +
                   (-11.25544*bp2_1) + (-8.38063*bp2_2) +
                   (-6.50522*bp2_3) + (-3.80130*bp2_4) + (-8.37399*gh1_1) +
                   (-5.56461*gh1_2) + (-3.02396*gh1_3) + (-1.31872*gh1_4) +
                   (-2.44706*vt2_1) + (-2.02168*vt2_2) + (-1.6185*vt2_3) +
                   (-1.14387*vt2_4) + (-0.42251*vt2_5) + (-0.33682*sf2_1) +
                   (-0.94342*sf2_2) + (-0.18043*sf2_3) + (0.11038*sf2_4) +
                   (3.04365*re2_1) + (2.32091*re3_1) + (3.46638*mh3_1) +
                   (2.90426*mh3_2) + (2.37241*mh3_3) + (1.36689*mh3_4) +
                   (0.66514*mh3_5) + (4.61446*mh4_1) + (3.41593*mh4_2) +
                   (2.34247*mh4_3) + (1.28044*mh4_4) + (0.41188*mh4_5),

           RAWMCS12 =
                   (3.93115*pf02_1) + (1.8684*pf02_2) +
                   (2.68282*pf04_1) + (1.43103*pf04_2) + (1.4406*rp2_1) +
                   (1.66968*rp3_1) + (1.48619*bp2_1) + (1.76691*bp2_2) +
                   (1.49384*bp2_3) + (0.90384*bp2_4) + (-1.71175*gh1_1) +
                   (-0.16891*gh1_2) + (0.03482*gh1_3) + (-0.06064*gh1_4) +
                   (-6.02409*vt2_1) + (-4.88962*vt2_2) + (-3.29805*vt2_3) +
                   (-1.65178*vt2_4) + (-0.92057*vt2_5) + (-6.29724*sf2_1) +
                   (-8.26066*sf2_2) + (-5.63286*sf2_3) + (-3.13896*sf2_4) +
                   (-6.82672*re2_1) + (-5.69921*re3_1) + (-10.19085*mh3_1) +
                   (-7.92717*mh3_2) + (-6.31121*mh3_3) + (-4.09842*mh3_4) +
                   (-1.94949*mh3_5) + (-16.15395*mh4_1) + (-10.77911*mh4_2) +
                   (-8.09914*mh4_3) + (-4.59055*mh4_4) + (-1.95934*mh4_5))
```

##### We then computed a norm-based standardization to PCS12 and MCS12.
```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
  PCS12 = (RAWPCS12 + 56.57706),
  MCS12 = (RAWMCS12 + 60.75781))
```

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
pss_1_coded =  as.numeric(pss_1_coded),
pss_2_coded =  as.numeric(pss_2_coded),
pss_3_coded = as.numeric(pss_3_coded),
pss_4_coded = as.numeric(pss_4_coded),
pss_5_coded = as.numeric(pss_5_coded))
```

##### For the SF-12 scale, a standardized SF-12 score was created for both physical and mental health for each participant. This resulted in the variables `PCS12` and `MCS12`. The higher the SF-12 scores, the better their overall health-related quality of life.

##### For the PSS scale, a total score was created from the questions which became the variable `total_PSS_score`. I then created stress categories for future categorical analysis: 
* Scores ranging from 0-6 would be considered low stress.
* Scores ranging from 7-13 ould be considered moderate stress.
* Scores ranging from 14-20 would be considered high perceived stress.

##### Further, several other variables were transformed. Age was transformed to a categorical variable, creating the age groups of 18-24, 25-44, 45-64, and 65+. Gender was transformed into a binary variable (0 = male, 1 = female), and a Housing Satisfaction Score was created. A higher score indicates higher overall housing satisfaction. This score was created based off of the totals of four questions: 
* The participant's satisfaction with their apartment
* The participant's satisfcation with their building 
* The participant's satisfcation with their neighbhorhood
* The participant's satisfaction with their property management 
* The participant's satisfcation with their apartment's temperature in the summer
* The participant's satisfaction with their apartment's temperature in the winter

##### Lastly, a binary chronic disease outcome was created. If a participant responded that they or someone in their home have the following conditions, they were coded as `1`: 
* Cancer
* Heart Disease
* Diabetes
* Hypertension
* Asthma
* Lung Disease 
* Anxiety
* Depression 
```{r}
SoBROData = 
SoBROData %>% 
  mutate(
    total_PSS_score = pss_1_coded + pss_2_coded + pss_3_coded + pss_4_coded + pss_5_coded,
    PSS_Category = total_PSS_score, 
    PSS_Category = 
          case_when(PSS_Category >= 14 ~ 'High Perceived Stress',
                    PSS_Category >= 7 ~ 'Moderate Stress',
                    PSS_Category >= 0  ~ 'Low Stress'),
    Age_Category = age,
    Age_Category = 
          case_when(Age_Category >= 65  ~ '65+',
                    Age_Category >= 45  ~ '45-64',
                    Age_Category >= 25 ~ '25-44',
                    Age_Category >= 18 ~ '18-24'),
    Gender_Coded = gender, 
    Gender_Coded = recode(Gender_Coded, 
                          "Male" = "0",
                          "Female" = "1",
                          `Other, please specify:` = "Other"),
    AptSat_Coded = aptsat, 
    AptSat_Coded = recode(AptSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"), 
    BuildingSat_Coded = buildsat, 
    BuildingSat_Coded = recode(BuildingSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
    NeighborhoodSat_Coded = neighsat, 
    NeighborhoodSat_Coded = recode(NeighborhoodSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   PropertyManSat_Coded = propertymansat,
   PropertyManSat_Coded = recode(PropertyManSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   WinTempSat_Coded = wintempsat, 
   WinTempSat_Coded = recode(WinTempSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"), 
   SumTempSat_Coded = sumtempsat, 
   SumTempSat_Coded = recode(SumTempSat_Coded,
                          "Very Dissatisfied" = "1",
                          "Dissatisfied" = "2", 
                          "Satisfied" = "3",
                          "Very Satisfied" ="4"),
   WinTempSat_Coded = as.numeric(WinTempSat_Coded),
   SumTempSat_Coded = as.numeric(SumTempSat_Coded),
   AptSat_Coded = as.numeric(AptSat_Coded), 
   BuildingSat_Coded = as.numeric(BuildingSat_Coded),
   NeighborhoodSat_Coded = as.numeric(NeighborhoodSat_Coded),
   PropertyManSat_Coded = as.numeric(PropertyManSat_Coded))

```

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
  PropertyManSat_Coded = na_mean(PropertyManSat_Coded),
  WinTempSat_Coded = na_mean(WinTempSat_Coded),
  HousingSatisfactionScore = AptSat_Coded + BuildingSat_Coded + NeighborhoodSat_Coded + PropertyManSat_Coded + WinTempSat_Coded + SumTempSat_Coded,
  Satisfaction_Category = HousingSatisfactionScore,
  Satisfaction_Category = 
          case_when(Satisfaction_Category >= 18 ~ 'High Satisfaction',
                    Satisfaction_Category >= 12 ~ 'Moderate Satisfaction',
                    Satisfaction_Category >= 6  ~ 'Low Satisfaction'))
```

##### Below, I changed variables into factors and numeric values.

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
    education = as.factor(education),
    Gender_Coded = factor(Gender_Coded, levels = c("0","1")),
    race = as.factor(race),
    age = as.numeric(age),
    income = factor(income, levels = c("1", "2", "3", "4", "5", "6","7","8","9","10")),
    smoking_status = factor(smoking_status, levels = c("0", "1")),
    micerats = factor(micerats,levels = c("0","1")),
    mildew = factor(mildew,levels = c("0","1")),
    cockroach = factor(cockroach,levels = c("0","1")),
    chronic_disease = factor(chronic_disease,levels = c("0","1")),
    hiv = factor(hiv,levels = c("0","1")))
```

##### In order to collapse education categories into fewer groups, education levels were finalized into three categories: 
* Less than high school (included those who never attended school, or did not complete a high school degree/G.E.D)
* High school/G.E.D
* Post High School Education (included those who attended vocational school, some college, 2-year/4-year college degree, and post graduate education)
```{r}
SoBROData = 
SoBROData %>% 
  mutate(education = as.character(education))
```

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
  Education_Group = education, 
  Education_Group = recode(Education_Group, 
                    "Never Attended School" = "Less than High School",
                    "Vocational School" = "Post High School Education",
                    "Some College" = "Post High School Education",
                    "2 Year Community College Degree" = "Post High School Education",
                    "4 Year College Degree" = "Post High School Education",
                    "Post-Graduate Degree" = "Post High School Education",
                    "High School/Secondary School Diploma" = 	"High School/G.E.D.",
                    "G.E.D." = "High School/G.E.D."))
```

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(Education_Group = as.factor(Education_Group))
```

##### In order to collapse income categories into fewer groups, education levels were finalized into three categories:
* `1` - Total personal income less than $10,000/year
* `2` - Total personal income between $10,000 - $24,000/year
* `3` - Total personal income $25,000k + /year

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
  Income_Group = income, 
  Income_Group = recode(Income_Group, 
                    "2" = "1",
                    "3" = "1",
                    "4" = "2",
                    "5" = "2",
                    "6" = "3",
                    "7" = "3",
                    "8" = "3",
                    "9" = "3",
                    "10" = "3"),
  Income_Group = factor(Income_Group,levels= c("1","2","3")))
```

#### In order to collapse racial categories into fewer groups, race/ethnicity was finalized into three categories: 
* Hispanic/Latinx
* Non-Hispanic Black or African American 
* Other (Included bi-racial/multiracial, Native American/Alaskan Native, and non-hispanic White)

```{r}
SoBROData = 
  SoBROData %>% 
  mutate(
  Racial_Group = race,
  Racial_Group = recode(Racial_Group,
                "Native American/ Alaskan Native" = "Other",
                "Non- Hispanic Black or African American,Hispanic or Latino,Asian or Pacific Islander,Other" = "Other",
                "Non- Hispanic Black or African American,Hispanic or Latino" = "Other",
                "Non- Hispanic Black or African American,Other, specify:" = "Other",
                "Non-Hispanic White" = "Other",
                "Non-Hispanic White,Hispanic or Latino" = "Other",
                "Non-Hispanic White,Non- Hispanic Black or African American,Hispanic or Latino,Native American/ Alaskan Native,Asian or Pacific Islander" = "Other",
                "Non-Hispanic White,Non- Hispanic Black or African American,Native American/ Alaskan Native,Other, specify:" = "Other",
                "Non-Hispanic White,Non- Hispanic Black or African American,Other, specify:" = "Other",
                
                "Non- Hispanic Black or African American,Hispanic or Latino,Asian or Pacific Islander,Other, specify:" = "Other",
"Non- Hispanic Black or African American,Hispanic or Latino,Other, specify:" = "Other"
                ))

```

### Univariable/Descriptive Analyses
##### The tables below provide descriptive analyses of the independent and dependent variables. 

#### Outcome Variables 
##### Overall Physical Health Status (Outcome 1)
```{r}
SoBROData %>% 
  select(PCS12) %>% 
  summary() %>% 
  knitr::kable(col.names = c("Summary of Self-Reported Physical Health"))
```

```{r}
SoBROData %>% 
  group_by(PCS12, Age_Category, gender) %>% 
  summarize(n = n()) %>% 
    ggplot(aes(x= Age_Category, y = PCS12)) + geom_boxplot() +
  geom_jitter() +  facet_grid(. ~ gender)
```

##### Overall Mentall Health Status (Outcome 2)
```{r}
SoBROData %>% 
  select(MCS12) %>% 
  summary() %>% 
  knitr::kable(col.names = c("Summary of Self-Reported Mental Health"))
```

```{r}
SoBROData %>% 
  group_by(MCS12, Age_Category, gender) %>% 
  summarize(n = n()) %>% 
    ggplot(aes(x= Age_Category, y = MCS12)) + geom_boxplot() +
  geom_jitter() +  facet_grid(. ~ gender)
```

#### Independent Variables
```{r}
SoBROData %>% 
  group_by(smoking_status) %>% 
  summarize(n=n()) %>% 
  mutate(
    percent = n/124*100) %>% 
  mutate(
    smoking_status = recode(smoking_status,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  knitr::kable(col.names=c("Smoking Status", "n", "Percent(%)"), digits = 2)
```

```{r}
SoBROData %>% 
  group_by(Racial_Group) %>% 
  summarize(n=n()) %>%
   mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Race/Ethnicity", "n", "Percent(%)"), digits = 2)
```

```{r}
SoBROData %>% 
  group_by(Age_Category) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Age Group", "n", "Percent(%)"), digits = 2)
```

```{r}
SoBROData %>% 
  group_by(gender) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Gender", "n", "Percent(%)"), digits = 2)
```


```{r}
SoBROData %>% 
  group_by(age, gender) %>% 
  ggplot(aes(x=gender, y=age)) + geom_violin() +
  labs(x = "Gender",
      y = "Age",
      title = "Figure 1. Distribution of Age by Gender")
```

```{r}
SoBROData %>% 
  drop_na() %>% 
  select(gender, age) %>% 
  group_by(gender) %>% 
  summarize(Mean = mean(age)) %>% 
  knitr::kable(col.names=c("Gender", "Mean Age"), digits = 2)
```

```{r}
SoBROData %>% 
  group_by(PSS_Category) %>% 
  summarize(n=n()) %>%
   mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Perceived Stress Score", "n", "Percent(%)"), digits = 2)

```

```{r}
SoBROData %>% 
  group_by(Income_Group) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Income Category", "n", "Percent(%)"), digits = 2)
```

```{r}
SoBROData %>% 
  group_by(education) %>% 
  summarize(n=n()) %>%
  mutate(
  percent = n/124*100) %>% 
  knitr::kable(col.names=c("Education Completed", "n", "Percent(%)"), digits = 2)
```

### 2x2 Tables and Visualizations
```{r}
SoBROData %>% 
  group_by(Age_Category, gender) %>% 
  summarize(n=n()) %>%
  pivot_wider(
    names_from = gender,
    values_from = n) %>% 
  knitr::kable(col.names = c("Age Group","Female (n)", "Male (n)", "Other (n)"))
```

```{r}
SoBROData %>%
  group_by(Age_Category, Racial_Group) %>% 
  summarize(n=n()) %>%
  pivot_wider(
    names_from = Racial_Group,
    values_from = n
  ) %>% 
  knitr::kable(col.names = c("Age Group", "Hispanic or Latinx (n)", "Other (n)", "Black/African American (n)", "NA"))
```

```{r}
SoBROData %>%
  group_by(Racial_Group, smoking_status) %>% 
  summarize(n=n()) %>%
  mutate(
    smoking_status = recode(smoking_status,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  pivot_wider(
    names_from = Racial_Group,
    values_from = n) %>% 
  knitr::kable(col.names = c("Smoking Status", "Hispanic or Latinx (n)", "Other (n)", "Black/African American (n)", "N/A"))
```

```{r}
SoBROData %>%
  group_by(Age_Category, smoking_status) %>% 
  summarize(n=n()) %>%
  mutate(
    smoking_status = recode(smoking_status,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")) %>% 
  pivot_wider(
    names_from = Age_Category,
    values_from = n) %>% 
  knitr::kable(col.names = c("Smoking Status", "18-24 (n)", "25-44 (n)", "45-64 (n)", "65+ (n)", "N/A"))
```

```{r}
SoBROData %>% 
  mutate(
    smoking_status = recode(smoking_status,
                   "0" = "Non-smoker(s)",
                   "1" = "Smoker(s)")
  ) %>% 
ggplot(aes(x=smoking_status, y=age)) + 
  geom_boxplot(outlier.shape=NA) + #avoid plotting outliers twice
  geom_jitter(position=position_jitter(width=.1, height=0)) + 
  labs(x = "Smoking Status",
      y = "Age",
      title = "Figure 4. Distribution of Age by Smoking Status")
```

```{r}
SoBROData %>% 
  group_by(Age_Category, Education_Group) %>% 
  ggplot(aes(x= Age_Category, fill = Education_Group)) + geom_bar() + coord_flip() +
  labs(x = "Age Group",
      y = "Count",
      color = "Education",
      title = "Figure 5. Education by Age Group") +
theme(legend.position = "bottom")
```
